Sequential rank agreement methods for comparison of ranked lists

### 2016-06-16 latex/superranker-manuscript.tex
Changes

With fully observed lists we can plot the sequential rank agreement
\ref{def:sra} as a function of depth $d$. An example is seen in top
panels of Figure~\ref{fig:example1} where four analysis methods were
used to rank 3051 gene expression values measured on 38 tumor mRNA
samples in order to classify between acute lymphoblastic leukemia
(ALL) and acute myeloid leukemia (AML)
\citep{Golub1999}. Preprocessing of the gene expression data was done
as described in \citet{Dudoit2002} and the four different analysis
approaches were: marginal two-sample $t$ tests, marginal logistic
regression analyses, elastic net logistic regression
\citep{friedman2010regularization}, and marginal maximum information
content correlations (MIC) \citep{Reshef2011}. For the first two
methods, the genes were ranked according to minimum $p$ value, for
logistic regression the genes were ordered by size of the
corresponding coefficients (after standardization), and MIC was
ordered by absolute correlation which resulted in the top rankings
seen in Table~\ref{tab1}.

The sequential rank agreement plotted in Figure~\ref{fig:example1}
shows how many ranks apart we on average expect the genes found in the
top $d$ part of the lists to be for each depth $d$.  It can be seen
that the sequential rank agreement limits are better towards the top
of the lists (smaller values on the $y$ axis) corresponding to
\emph{better} agreement than in the tails of the lists. It is clear
from the curve in Figure~\ref{fig:example1} that there is a
substantial drop in agreement even after the first depth. Thus, if we
were to restrict attention to a small set of predictors then we would
focus on predictors 2124, 829, and 378 as seen in Table~\ref{tab1}.

@@

\begin{figure}[htbp]
   \begin{center}
 \includegraphics[width=.49\textwidth]{paper-fig1a}
 \includegraphics[width=.49\textwidth]{paper-fig2a}
 \includegraphics[width=.49\textwidth]{paper-fig1b}
 \includegraphics[width=.49\textwidth]{paper-fig2b}
 \end{center}
 \caption{Left panels: Sequential rank agreement for 4 different
   analysis methods applied to the 3051 genes in the Golub data /black
   line). Right
   panels: Corresponding sequential rank agreement for the same data
   but where only the top 20 ranked items are available. The blue and
   red areas correspond to the independent and randomized reference
   hypothesis areas, respectively. The bottom plots are identical to
   the corresponding top plots but have been zoomed in on the first
   part of the $x$ axis.}
 \label{fig:example1}
 \end{figure}

 If the lists agree on all items then the sequential rank agreement is
 zero for all depths. Generally, if there are changes in the level of
 rank agreement then this suggests that there are sets of items that
 are ranked similarly in all lists and other sets of items that have
 been assigned vastly different ranks in the lists. In gene
 association studies we generally expect the agreement among the lists
 to be better towards the top of the lists and worse towards the end
 of the lists. In this case the sequential rank agreement curve will
 start at a low level and then increase until it levels off as in
 Figure~\ref{fig:example1}.

\subsection{Analysis of partial/censored lists}
Incomplete or partial lists are also a common occurrence. They arise
for example in case of missing data (items), when comparing top $d$
list results from publications, or when some methods only rank a
subset of the items. For example, penalized regression based on the
Lasso provides a sparse set of predictors that have non-zero
coefficients. There is no obvious ordering of the set of predictors
whose coefficient has been shrunk to zero and thus we end up with a
partial ordering. Incomplete lists also occur when the analyst decides
to censor all lists at depth $d$ such that the measurement of
agreement is restricted to the first $d$ elements of each list, or
when the analyst only wants to consider the ranking of items that have
been found to be statistically significant.

@@

The agreement, $A(X_p)$, cannot be computed directly for all
predictors in the presence of censored lists because the exact rank
for some items will be unknown. Also, recall that the rankings within
a single list are not independent since each rank must appear exactly
once in each list. Thus, we cannot simply assign the same number
(e.g., the mean of the unassigned ranks) to the censored items since
that would result in less variation of the ranks and hence less
variation of the agreement, and it would artificially introduce a
(downward) bias of agreement for items that are censored in multiple
lists.

@@

\subsection{Comparison to average overlap}
The average overlap is a measure that is widely used in comparisons of
two fully observed ranked lists \citep{Fagin2003,Webber2010} and the
idea behind the average overlap closely resembles the sequential rank
agreement. In this section we present an extension of the average
overlap to more that two lists and compare this to sequential rank
agreement.

The overlap among $L$ lists observed until depth $d$ is defined as the
proportion of items that are found in all $L$ lists compared to the
possible number of items found in all lists observed until depth $d$:
\begin{align}
  O(d) = \frac{| \cap_{l=1}^L \{R_l^{-1}(r); r\leq d \} |}{d}\label{def:overlap}
\end{align}
Typically the overlap is only considered for pairwise comparisons
\citep{Bar-Ilan2006,Boulesteix2009} but formula \ref{def:overlap}
is easily extended to the situation where $L>2$ as shown above.

The average overlap at depth $d$ is defined as the average of the
overlaps until depth $d$,
\begin{equation}
AO(d) = \frac1d\sum_{i=1}^d O(d).
\end{equation}
This definition directly ensures that further emphasis is put on items
in the top of the lists since their item specific overlap contributes
to the calculation of average overlap at higher depths.

There are three main drawbacks of the average overlap method: First it
is highly sensitive to single items in the top of one the lists.  The
average overlap profile may change dramatically depending on the
presence or absence of similar items in the beginning of the
lists. For example, if all lists have the same item at the top then
the average overlap attains its maximum at 1, but if just one of the
lists differ then the overage overlap starts at its minimum of
0. Secondly, when the number of lists, $L$, is high then it might be
difficult to obtain a non-zero overlap (and hence a non-zero average
overlap) at the beginning of the lists because the overlap will be
zero until an item is present in all $L$ lists. Finally, the average
overlap has an interpretation in terms of moving averages of
percentages (of $L$ sets) which is somewhat less intuitive than our
proposed sequential rank distance.

In the following we compare the average overlap and the sequential
rank agreement using the analysis of the Golub data
\citep{Golub1999}. We consider a typical application of the average
overlap and thus restrict attention to only two gene rankings obtained
by the $t$ tests and logistic regression. These two marginal modeling
approaches should result in very similar gene rankings. The first two
columns of Table~\ref{tab1} show the actual top-10 ranking of the
predictors.

% chunks 8,9

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=.49\textwidth]{paper-fig-ao1}
\includegraphics[width=.49\textwidth]{paper-fig-ao2}
\end{center}
 \caption{Average overlap (left figure) and sequential rank agreement
 (right figure) for comparing two different analysis methods
 (marginal $t$ test and logistic regression)
    applied to the Golub data. The black lines are based on the full
    set of predictors while the red lines have the two highest
    associated predictors removed
    from the data before computing the average overlap and sequential
    rank agreement.}
  \label{fig:case1}
 \end{figure}

 The black lines in Figure~\ref{fig:case1} show the average overlap
 and sequential rank agreement for these data. It is clear from both
 plots that there is perfect agreement towards the top of the ranked
 lists: the average overlap is 1 and the sequential rank agreement is
 zero.  However, if we remove those two items from the data (i.e.,
 gene/predictor 2124 and 896) and redo the analyses then we get
 substantially different curves for the average overlap but not for
 sequential rank agreement (red lines in Figure~\ref{fig:case1}). For
 the sequential rank agreement we get roughly the same estimate of
 agreement from rank 3 as we did from the full dataset. This indicates
 that sequential rank is more robust against small perturbations of
 the data. For the average overlap the curves completely change when
 these two items are removed and which would lead to a different
 conclusion about the agreement among the items from rank 3 downwards.

@@

The first reference hypothesis is
\begin{eqnarray}
H_0  & : &  \text{The list rankings correspond to complete randomly}\label{eq:permutationHypothesis}\\
       &  & \text{permuted lists}\nn
\end{eqnarray}
which not only assumes that there is no information in the data on
which the rankings are based but also that the methods are completely
independent.

Alternatively, we can remove the restriction on the independence among
the methods by only requiring that there is no information in the
ranking but that the rankings are all based on applying the
method/approaches to the same data
\begin{eqnarray*}
\widetilde H_0 & :&  \text{The list rankings are based on data containing}\\
& &   \text{no association to the outcome.}
\end{eqnarray*}
$H_0$ is a quite unrealistic null hypothesis but we can easily obtain
realizations from that null hypothesis simply by permuting the items
within each list and then computing the sequential rank agreement for
the permuted lists. In the fully observed case each experiment
contains $L$ lists of random permutations of the items in $X$. For the
censored/partially observed case we first permute the items
$X_1,\dots,X_P$ and then censor list $l$ at $d_l$. The sequential rank
agreement curve from the the original lists can then be compared to,
say, the pointwise 95\% quantiles of the observed rank agreements
obtained under $H_0$.

@@

\section{Asymptotics}\label{sec:chgpoint}
In many applications it is of interest to estimate a list depths which
satisfies a change point criterion. In this section we investigate the
theoretical properties of our proposed method for this specific
task. As in \cite{hall:schi:2012} we work with an infinite set
of lists and study the asymptotic behaviour for $L\to\infty$.
However, we do not allow the length of the lists to change with $L$
because in most applications the lengths of the lists are fixed.

@@

\begin{theorem}
  Assume that $\{R_l(X)\}_{l=1}^L$ are independent draws from the a
  probability distribution $Q$ on the set of lists.  Then,
  $\left\|\widehat{\textrm{sra}}_L - \textrm{sra}\right\|_\infty =
  o_P(1)$.
\label{thm:consistency}
\end{theorem}

We now define the parameter of interest. For any given deterministic threshold function $q\colon\, 1,\ldots,P \mapsto \mathbb{R}_{\geq 0}$
we define the population superlevel set of the sequential rank agreement as
\begin{align}
  \mathcal{L}(q) = \left\{d : \textrm{sra}(d) \geq q(d)\right\}
\end{align}
The values of $q$ could for example be the limits-of-agreement for randomly
permuted lists as in the null-hypothesis in equation (\ref{eq:permutationHypothesis}).
A change-point $d^\ast(q)$ in the list agreement is then defined by the position
\begin{align}
  d^\ast(q) &= 1\left(|\mathcal{L}(q)| > 0\right)\inf \mathcal{L}(q) + 1\left(|\mathcal{L}(q)| = 0\right)P
\end{align}
corresponding to the first list depth where the sequential rank agreement exceeds the threshold 
if such a position exists. Otherwise, the full list is in agreement according to $q$ and the change-point
is defined as the length of the lists. The empirical superlevel set is similarly defined as
\begin{align}
 \widehat{\mathcal{L}}_L(\widehat{q}_L) = \left\{d : \widehat{\textrm{sra}}_L(d) \geq \widehat{q}_L(d)\right\}
\end{align}
where we allow that the threshold function can depend on the sample size as well. The estimated change-point is therefore
\begin{align}
  \widehat{d^\ast_L}(\widehat{q}_L) &= 1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| > 0)\inf \widehat{\mathcal{L}}_L(\widehat{q}_L) + 1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| = 0)P
\end{align}

@@

\mcomment{Elaborate more on how to apply this in practice (finite sample/bootstrap/etc.) and that $q$ is similar to LoA BA-style?}

@@

The sequential rank agreement given in equation
(\ref{def:sra}) may therefore be rewritten as the following integral
with respect to $N$
\begin{align}
  \textrm{sra}(d) = \int_{1}^P \frac{A(X_p)}{N(P; d)}\mathrm dN(p; d).
\end{align}
The empirical sequential rank agreement is similarly given by
\begin{align}
  \widehat{\textrm{sra}}_L(d) &= \int_1^P \frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)}\mathrm d \widehat{N}_L(p; d) 
\end{align}
and it follows that
\begin{align}
  \sup_{d \in 1,\ldots,P}\left|\widehat{\textrm{sra}}_L(d) - \textrm{sra}(d)\right| \leq \sup_{d \in 1,\ldots,P}\left|U_L(d)\right| - \sup_{d \in 1,\ldots,P}\left|V_L(d)\right|
\end{align}
where
\begin{align}
  U_L(d) &= \int_1^P\frac{A(X_p)}{N(P;d)} \mathrm d\left(\widehat{N}_L(p; d) - N(p; d)\right)\\
  V_L(d) &=  \int_1^P \left(\frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)} - \frac{A(X_p)}{N(P;d)}\right)\mathrm d\widehat{N}_L(p; d)
\end{align}
The conclusion of the proof therefore follows if each of these two
terms are uniformly of order $o_P(1)$ in $d$.  For the first term we
have that
\begin{align}
  \left|U_L(d)\right| \leq \left(\sup_{p \in 1,\ldots,P}\left|\frac{A(X_p)}{N(P;d)}\right|\right)\left(\widehat{N}_L(p; d) - N(p; d)\right)\biggr\rvert_1^P
\end{align}
where the second factor is $o_P(1)$ uniformly in $d$ by equation
(\ref{eq:countingConv}). The first factor is $O(1)$ since $N(P;d)>0$
uniformly in $d$ and because $A(X_p) = O(P^2)$ uniformly in $p$ by the Cauchy-Ekstr{\o}m inequality. It
thus follows that $\left|U_L(d)\right| = o_P(1)$ uniformly in $d$.

>>>>

With fully observed lists we can plot the sequential rank agreement
\eqref{def:sra} as a function of depth $d$. An example is seen in top
panels of Figure~\ref{fig:example1} where four analysis methods were
used to rank 3051 gene expression values measured on 38 tumor mRNA
samples in order to classify between acute lymphoblastic leukemia
(ALL) and acute myeloid leukemia (AML)
\citep{Golub1999}. Preprocessing of the gene expression data was done
as described in \citet{Dudoit2002} and the four different analysis
approaches were: marginal unequal variances two-sample $t$ tests,
marginal logistic regression analyses, elastic net logistic regression
\citep{friedman2010regularization}, and marginal maximum information
content correlations (MIC) \citep{Reshef2011}. For the first two
methods, the genes were ranked according to minimum $p$ value, for
logistic regression the genes were ordered by size of the
corresponding coefficients (after standardization), and MIC was
ordered by absolute correlation which resulted in the top rankings
seen in Table~\ref{tab1}.

The sequential rank agreement curve plotted in Figure~\ref{fig:example1}
roughly shows how many ranks apart we on average expect the genes found in the
top $d$ part of the lists to be for each depth $d$.  It can be seen
that the sequential rank agreement limits are better towards the top
of the lists (smaller values on the $y$ axis) corresponding to
\emph{better} agreement than towards the tails of the lists. It is clear
from the curve in Figure~\ref{fig:example1} that there is a
substantial drop in agreement even after the first depth. Thus, if we
were to restrict attention to a small set of predictors then we would
focus on predictors 2124, 829, and 378 as seen in Table~\ref{tab1}.

@@

\begin{figure}[htbp]
   \begin{center}
 \includegraphics[width=.49\textwidth]{paper-fig1a}
 \includegraphics[width=.49\textwidth]{paper-fig2a}
 \includegraphics[width=.49\textwidth]{paper-fig1b}
 \includegraphics[width=.49\textwidth]{paper-fig2b}
 \end{center}
 \caption{Left panels: Sequential rank agreement for 4 different
   analysis methods applied to the 3051 genes in the Golub data /black
   line). Right
   panels: Corresponding sequential rank agreement for the same data
   but where only the top 20 ranked items are available and the rest
   of the items have been censored. The blue and
   red areas correspond to the independent and randomized reference
   hypothesis areas, respectively. The bottom plots are identical to
   the corresponding top plots but have been zoomed in on the first
   part of the $x$ axis.}
 \label{fig:example1}
 \end{figure}

 If the lists agree on all items then the sequential rank agreement is
 zero for all depths. Generally, if there are changes in the level of
 rank agreement then this suggests that there are sets of items that
 are ranked similarly in all lists and other sets of items that have
 been assigned vastly different ranks in the lists. In gene
 association studies we generally expect the agreement among the lists
 to be better towards the top of the lists and worse towards the end
 of the lists. In this case the sequential rank agreement curve will
 start at a low level and then increase until it levels off as seen in
 Figure~\ref{fig:example1}.

\subsection{Analysis of partial/censored lists}
Incomplete or partial lists are a common occurrence that arise, for
example, in case of missing data (items), when comparing top $d$ list
results from publications, or when some methods only rank a subset of
the items. For example, penalized regression based on the Lasso
provides a sparse set of predictors that have non-zero
coefficients. There is no obvious ordering of the set of predictors
whose coefficient has been shrunk to zero and thus we end up with a
partial ordering. Incomplete lists also occur when the analyst decides
to censor all lists at depth $d$ (\ie, top $d$ lists) such that the
measurement of agreement is restricted to the first $d$ elements of
each list, or when the analyst only wants to consider the ranking of
items that have been found to be statistically significant.

@@

The agreement, $A(X_p)$, cannot be computed directly for all
predictors in the presence of censored lists because the exact rank
for some items will be unknown. Also, recall that the rankings within
a single list are not independent since each rank must appear exactly
once in each list. Thus, we cannot simply assign the same number
(e.g., the mean of the unassigned ranks) to the censored items since
that would result in \emph{less} variation of the ranks and hence less
variation of the agreement, and it would artificially introduce a
(downward) bias of agreement for items that are censored in multiple
lists.

@@

% \subsection{Comparison to average overlap}
% The average overlap is a measure that is widely used in comparisons of
% two fully observed ranked lists \citep{Fagin2003,Webber2010} and the
% idea behind the average overlap closely resembles the sequential rank
% agreement. In this section we present an extension of the average
% overlap to more that two lists and compare this to sequential rank
% agreement.

% The overlap among $L$ lists observed until depth $d$ is defined as the
% proportion of items that are found in all $L$ lists compared to the
% possible number of items found in all lists observed until depth $d$:
% \begin{align}
%   O(d) = \frac{| \cap_{l=1}^L \{R_l^{-1}(r); r\leq d \} |}{d}\label{def:overlap}
% \end{align}
% Typically the overlap is only considered for pairwise comparisons
% \citep{Bar-Ilan2006,Boulesteix2009} but formula \ref{def:overlap}
% is easily extended to the situation where $L>2$ as shown above.

% The average overlap at depth $d$ is defined as the average of the
% overlaps until depth $d$,
% \begin{equation}
% AO(d) = \frac1d\sum_{i=1}^d O(d).
% \end{equation}
% This definition directly ensures that further emphasis is put on items
% in the top of the lists since their item specific overlap contributes
% to the calculation of average overlap at higher depths.

% There are three main drawbacks of the average overlap method: First it
% is highly sensitive to single items in the top of one the lists.  The
% average overlap profile may change dramatically depending on the
% presence or absence of similar items in the beginning of the
% lists. For example, if all lists have the same item at the top then
% the average overlap attains its maximum at 1, but if just one of the
% lists differ then the overage overlap starts at its minimum of
% 0. Secondly, when the number of lists, $L$, is high then it might be
% difficult to obtain a non-zero overlap (and hence a non-zero average
% overlap) at the beginning of the lists because the overlap will be
% zero until an item is present in all $L$ lists. Finally, the average
% overlap has an interpretation in terms of moving averages of
% percentages (of $L$ sets) which is somewhat less intuitive than our
% proposed sequential rank distance.

% In the following we compare the average overlap and the sequential
% rank agreement using the analysis of the Golub data
% \citep{Golub1999}. We consider a typical application of the average
% overlap and thus restrict attention to only two gene rankings obtained
% by the $t$ tests and logistic regression. These two marginal modeling
% approaches should result in very similar gene rankings. The first two
% columns of Table~\ref{tab1} show the actual top-10 ranking of the
% predictors.

% % chunks 8,9

% \begin{figure}[htbp]
% \begin{center}
% \includegraphics[width=.49\textwidth]{paper-fig-ao1}
% \includegraphics[width=.49\textwidth]{paper-fig-ao2}
% \end{center}
%  \caption{Average overlap (left figure) and sequential rank agreement
%  (right figure) for comparing two different analysis methods
%  (marginal $t$ test and logistic regression)
%     applied to the Golub data. The black lines are based on the full
%     set of predictors while the red lines have the two highest
%     associated predictors removed
%     from the data before computing the average overlap and sequential
%     rank agreement.}
%   \label{fig:case1}
%  \end{figure}

%  The black lines in Figure~\ref{fig:case1} show the average overlap
%  and sequential rank agreement for these data. It is clear from both
%  plots that there is perfect agreement towards the top of the ranked
%  lists: the average overlap is 1 and the sequential rank agreement is
%  zero.  However, if we remove those two items from the data (i.e.,
%  gene/predictor 2124 and 896) and redo the analyses then we get
%  substantially different curves for the average overlap but not for
%  sequential rank agreement (red lines in Figure~\ref{fig:case1}). For
%  the sequential rank agreement we get roughly the same estimate of
%  agreement from rank 3 as we did from the full dataset. This indicates
%  that sequential rank is more robust against small perturbations of
%  the data. For the average overlap the curves completely change when
%  these two items are removed and which would lead to a different
%  conclusion about the agreement among the items from rank 3 downwards.

@@

The first reference hypothesis is
\begin{eqnarray}
H_0  & : &  \text{The list rankings correspond to completely randomly}\label{eq:permutationHypothesis}\\
       &  & \text{permuted lists}\nn
\end{eqnarray}
which not only assumes that there is no information in the data on
which the rankings are based but also that the methods are completely
independent.

Alternatively, we can remove the restriction on the independence among
the methods by only requiring that there is no information in the
ranking but that the rankings are all based on applying the
method/approaches to the same data
\begin{eqnarray*}
\widetilde H_0 & :&  \text{The list rankings are based on data containing}\\
& &   \text{no association to the outcome.}
\end{eqnarray*}

\subsection{Permutation-based inference}

$H_0$ is a quite unrealistic null hypothesis but we can easily obtain
realizations from that null hypothesis simply by permuting the items
within each list and then computing the sequential rank agreement for
the permuted lists. In the fully observed case each experiment
contains $L$ lists of random permutations of the items in $X$. For the
censored/partially observed case we first permute the items
$X_1,\dots,X_P$ and then censor list $l$ at $d_l$ (\ie, each list so
it has the same number of observed rankings as was observed for list
$l$ in the original dataset). The sequential rank agreement curve from
the the original lists can then be compared to, say, the pointwise
95\% quantiles of the observed rank agreements obtained under $H_0$.

@@

\subsection{Asymptotic inference of change in agreement}\label{sec:chgpoint}
In many applications it is of interest to estimate a list depth which
satisfies a change point criterion since that corresponds to a change
in agreement among the list ranks. In this section we investigate the
theoretical properties of our proposed method for this specific
task. As in \cite{hall:schi:2012} we consider an infinite set
of lists and study the asymptotic behaviour for $L\to\infty$.
However, we do not allow the length of the lists to change with $L$
because in most applications the lengths of the lists are fixed.

@@

\begin{theorem}
  Assume that $\{R_l(X)\}_{l=1}^L$ are independent draws from a
  probability distribution $Q$ on the set of lists.  Then,
  $\left\|\widehat{\textrm{sra}}_L - \textrm{sra}\right\|_\infty =
  o_P(1)$.
\label{thm:consistency}
\end{theorem}

We now define the parameter of interest. For any given deterministic threshold function $q\colon\, 1,\ldots,P \mapsto \mathbb{R}_{\geq 0}$
we define the population superlevel set of the sequential rank agreement as
\begin{align}
  \mathcal{L}(q) = \left\{d : \textrm{sra}(d) \geq q(d)\right\}.
\end{align}
The values of $q$ could for example be the limits-of-agreement for randomly
permuted lists as in the null-hypothesis in equation (\ref{eq:permutationHypothesis}).
A change-point $d^\ast(q)$ in the list agreement is then defined by the position
\begin{align}
  d^\ast(q) &= 1\left(|\mathcal{L}(q)| > 0\right)\inf \mathcal{L}(q) + 1\left(|\mathcal{L}(q)| = 0\right)P
\end{align}
corresponding to the first list depth where the sequential rank agreement exceeds the threshold 
if such a position exists. Otherwise, the full list is in agreement according to $q$ and the change-point
is defined as the length of the lists. The empirical superlevel set is similarly defined as
\begin{align}
 \widehat{\mathcal{L}}_L(\widehat{q}_L) = \left\{d : \widehat{\textrm{sra}}_L(d) \geq \widehat{q}_L(d)\right\}
\end{align}
where we allow the threshold function to depend on the sample size as well. The estimated change-point is therefore
\begin{align}
  \widehat{d^\ast_L}(\widehat{q}_L) &= 1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| > 0)\inf \widehat{\mathcal{L}}_L(\widehat{q}_L) + 1(|\widehat{\mathcal{L}}_L(\widehat{q}_L)| = 0)P
\end{align}

@@

\mcomment{Elaborate more on how to apply this in practice (finite
  sample/bootstrap/etc.) and that $q$ is similar to LoA BA-style? Show
some result althrough }

@@

The sequential rank agreement given in equation
(\ref{def:sra}) may therefore be rewritten as the following integral
with respect to $N$
\begin{align}
  \textrm{sra}(d) = \int_{1}^P \frac{A(X_p)}{N(P; d)}\mathrm dN(p; d).
\end{align}
The empirical sequential rank agreement is similarly given by
\begin{align}
  \widehat{\textrm{sra}}_L(d) &= \int_1^P \frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)}\mathrm d \widehat{N}_L(p; d) 
\end{align}
and it follows that
\begin{align}
  \sup_{d \in 1,\ldots,P}\left|\widehat{\textrm{sra}}_L(d) - \textrm{sra}(d)\right| \leq \sup_{d \in 1,\ldots,P}\left|U_L(d)\right| - \sup_{d \in 1,\ldots,P}\left|V_L(d)\right|
\end{align}
where
\begin{align}
  U_L(d) &= \int_1^P\frac{A(X_p)}{N(P;d)} \mathrm d\left(\widehat{N}_L(p; d) - N(p; d)\right)\\
  V_L(d) &=  \int_1^P \left(\frac{\widehat{A}_L(X_p)}{\widehat{N}_L(P; d)} - \frac{A(X_p)}{N(P;d)}\right)\mathrm d\widehat{N}_L(p; d)
\end{align}
The conclusion of the proof therefore follows if each of these two
terms are uniformly of order $o_P(1)$ in $d$.  For the first term we
have that
\begin{align}
  \left|U_L(d)\right| \leq \left(\sup_{p \in 1,\ldots,P}\left|\frac{A(X_p)}{N(P;d)}\right|\right)\left(\widehat{N}_L(p; d) - N(p; d)\right)\biggr\rvert_1^P
\end{align}
where the second factor is $o_P(1)$ uniformly in $d$ by equation
(\ref{eq:countingConv}). The first factor is $O(1)$ since $N(P;d)>0$
uniformly in $d$ and because $A(X_p) = O(P^2)$ uniformly in $p$ by the Cauchy-Schwarz inequality. It
thus follows that $\left|U_L(d)\right| = o_P(1)$ uniformly in $d$.
